---
title: "DATA624_HW8"
author: "John Mazon"
date: "4/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyr)
library(forecast)
library(fpp)
library(fpp3)
library(fpp2)
library(glue)
library(caret)
library(caTools)
library(pls)
library(Amelia)
library(RANN)
library(mlbench)
library(caret)
```

## 7.2

7.2. Friedman (1991) introduced several benchmark data sets create by sim-ulation. One of these simulations used the following nonlinear equation to create data:
y = 10 sin(πx1x2) + 20(x3 − 0.5)2 + 10x4 + 5x5 + N (0, σ2)
where the x values are random variables uniformly distributed between [0, 1]
(there are also 5 other non-informative variables also created in the simula-
tion). The package mlbench contains a function called mlbench.friedman1 that
simulates these data:

```{r}
 set.seed(69)
 the_training_data <- mlbench.friedman1(200, sd = 1)
 ## here I convert the 'x' data from a matrix to a data frame

 the_training_data$x <- data.frame(the_training_data$x)
 
 featurePlot(the_training_data$x, the_training_data$y)


 ## This creates a list with a vector 'y' and a matrix
 ## of predictors 'x'. 
 ## Notice that we also simulate a large test set to estimate the true error rate with good precision:
 the_test_data <- mlbench.friedman1(5000, sd = 1)
 the_test_data$x <- data.frame(the_test_data$x)


```


Tune several models on these data. For example:

```{r}


the_knn_model <- train(x = the_training_data$x,
 y = the_training_data$y,
 method = "knn",

 preProc = c("center", "scale"),
 tuneLength = 10)
 the_knn_model
```


 
200 samples
10 predictors


```{r}
 my_knn_pred <- predict(the_knn_model, newdata = the_test_data$x)
## The function 'postResample' can be used to get the test set
 ## perforamnce values
 postResample(pred = my_knn_pred, obs = the_test_data$y)


```

Which models appear to give the best performance? Does MARS select the
informative predictors (those named X1–X5)?




```{r}
the_NnetGrid <- expand.grid(.decay = c(0,0.01,.1),
                        .size = c(1:5),
                        .bag = FALSE)

my_nnetFit <- train(the_training_data$x, the_training_data$y,
                  method = 'avNNet',
                  tuneGrid = the_NnetGrid,
                  preProc = c('center','scale'),
                  linout = TRUE,
                  trace = FALSE,
                  MaxNWts = 5 * (ncol(the_training_data$x) + 1 + 5 + 1),
                  maxit = 100
  
)

head(my_nnetFit)
```



```{r}
nnetPred <- predict(my_nnetFit, newdata = the_test_data$x)
postResample(pred = nnetPred, obs = the_test_data$y)

```


```{r}
#creating tune grid
marsGrid <- expand.grid(.degree = 1:2, .nprune = 2:28)
set.seed(100)
my_mars_tune <- train(the_training_data$x, the_training_data$y,
                   method = 'earth',
                   tuneGrid = marsGrid,
                   trControl = trainControl(method = 'cv'))

my_mars_tune

```

```{r}
varImp(my_mars_tune)
```


```{r}
my_mars_pred <- predict(my_mars_tune, newdata = the_test_data$x)
postResample(pred = my_mars_pred, obs = the_test_data$y)

```



```{r}
the_SVMR_tune <- train(the_training_data$x, the_training_data$y,
                   method = 'svmRadial',
                   preProc = c('center','scale'),
                   tuneLength = 14,
                   trControl = trainControl(method = 'cv'))

the_SVMR_tune

```

```{r}
my_SVM_pred <- predict(the_SVMR_tune, newdata = the_test_data$x)
postResample(pred = my_SVM_pred, obs = the_test_data$y)

my_knn_tune <- train(the_training_data$x, the_training_data$y,
                   method = 'knn',
                   preProc = c('center','scale'),
                   tuneGrid = data.frame(.k = 1:20),
                   trControl = trainControl(method = 'cv'))

my_knn_tune


my_knn_pred <- predict(my_knn_tune, newdata = the_test_data$x)
postResample(pred = my_knn_pred, obs = the_test_data$y)
```


We can see that our MARS model produces the greatest result. The Mars model only uses the informative predictors, X1-X5.

## 7.5

7.5. Exercise 6.3 describes data for a chemical manufacturing process. Use
the same data imputation, data splitting, and pre-processing steps as before
and train several nonlinear regression models.


```{r}
library(RANN)
library(AppliedPredictiveModeling)
data(ChemicalManufacturingProcess)
```

```{r}
(the_chem_imput <- preProcess(ChemicalManufacturingProcess[,-c(1)], method=c('knnImpute')))

the_chem_mod <- predict(the_chem_imput, ChemicalManufacturingProcess[,-c(1)])
remove_cols <- nearZeroVar(the_chem_mod, names = TRUE, 
                           freqCut = 2, uniqueCut = 20)
all_cols <- colnames(the_chem_mod)
the_chem_mod <- the_chem_mod[ , setdiff(all_cols,remove_cols)]

my_train_row <- sort(sample(nrow(the_chem_mod), nrow(the_chem_mod)*.7))
my_train_x <- the_chem_mod[my_train_row,]
my_testSET_x <- the_chem_mod[-my_train_row,]
my_testSET_y <- ChemicalManufacturingProcess[my_train_row,1]
test_y_set <- ChemicalManufacturingProcess[-my_train_row,1]
```

(a) Which nonlinear regression model gives the optimal resampling and test
set performance?

MARS
```{r}
my_MARS_mod <- earth(x = my_train_x,
                  y = my_testSET_y)
my_MARS_mod

```


SVM
```{r}
my_SVM_mod <- train(x = my_train_x,
                  y = my_testSET_y,
                  method = "svmRadial",
                  preProc = c("center", "scale"),
                  tuneLength = 10,
                  trcontrol = trainControl(method = "cv"))
my_SVM_mod
plot(my_MARS_mod, which = 1)
```


KNN
```{r}
the_knn_model <- train(x = my_train_x,
                  y = my_testSET_y,
                  method = "knn",
                  preProc = c("center", "scale"),
                  tuneLength = 10)
the_knn_model

```


As we can see when viewing the R2 values, the MARS model is the best test set performance. You'll notice that only 9 predictors are selected in this model.


(b) Which predictors are most important in the optimal nonlinear regres-
sion model? Do either the biological or process variables dominate the
list? How do the top ten important predictors compare to the top ten
predictors from the optimal linear model?
```{r}

evimp(my_MARS_mod)
```

If we look back at 6.3 our linear model from our results found Manufacturing Processes 20, 32, 6, 9, 13 and 36 as the most important.
Three of these predictors are also present in our MARS model from at the top.
Manufacturing processes are strongest on this list similar to that of our optimal linear model.



(c) Explore the relationships between the top predictors and the response for
the predictors that are unique to the optimal nonlinear regression model.
Do these plots reveal intuition about the biological or process predictors
and their relationship with yield?

```{r}

the_best_predi <- c("ManufacturingProcess32", "ManufacturingProcess09", "ManufacturingProcess13","ManufacturingProcess01", "ManufacturingProcess42","ManufacturingProcess43")
featurePlot(my_train_x[,the_best_predi], my_testSET_y)

```


Manufacturing processes 32 and 9 appear to have positive relationships with our predictor.
Then you'll notice that meanwhile Process 13 shows a similar negative relationship. 
Our relationships for Processes 1, 42 and 43 seem to be largely influenced by outliers in the data.
However though removing these outliers before the modeling led to significant drops in the accuracy of the models.
We can say that this suggests they are not necessarily erroneous data points.


